---
title: "Adni Preprocessing"
author: "Aborageh"
date: "8/17/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Users/maborageh/Desktop/Master Thesis/ADNIMERGE/data')
```

```{r}
#plyr has to be loaded before dplyr to avoid conflicts
library(plyr)
library(dplyr)
library(plotly)
library(qwraps2)
library(readr)
```

```{r}
load(paste(getwd(), "/adnimerge.rdata", sep = ""))
adnimerge$long.age <- round((adnimerge[,"AGE"] + adnimerge[,"Years.bl"]), digits =1)
```

```{r}
length(unique(adnimerge$RID))
```
### ADNI Features###
```{r}
adni.features <- adnimerge[,c("RID", "VISCODE","DX","PTGENDER","PTEDUCAT","PTETHCAT","PTRACCAT","PTMARRY","APOE4","AGE","Ventricles","Hippocampus","MidTemp","ICV","WholeBrain","Entorhinal","Fusiform")]
```

### FAQ Data ###
```{r}
load(paste(getwd(), "/faq.rdata", sep = ""))
faq[3966,5] <- "m156"
print("Number of Patients:")
length(unique(faq$RID))
```

### MMSE Features###
```{r}
load(paste(getwd(), "/mmse.rdata", sep = ""))
```

```{r}
mmse = mmse
#mmse[mmse=="Incorrect"]<-0
subItemScores = c("MMDATE", "MMYEAR", "MMMONTH", "MMDAY", "MMSEASON", "MMHOSPIT", "MMFLOOR", "MMCITY", "MMAREA", "MMSTATE", "MMBALL", "MMFLAG", "MMTREE", "MMD", "MML", "MMR", "MMO", "MMW", "MMBALLDL", "MMFLAGDL", "MMTREEDL", "MMWATCH", "MMPENCIL", "MMREPEAT", "MMHAND", "MMFOLD", "MMONFLR", "MMREAD", "MMWRITE", "MMDRAW")

 

mmse <- data.frame(lapply(mmse, function(x) {
                 gsub("0 - incorrect", 0, x)
            }))
mmse <- data.frame(lapply(mmse, function(x) {
                 gsub("0 - Incorrect",0, x)
            }))
mmse <- data.frame(lapply(mmse, function(x) {
                 gsub("1 - Correct", 1, x)
            }))
mmse <- data.frame(lapply(mmse, function(x) {
                 gsub("1 - correct", 1, x)
            }))
mmse <- data.frame(lapply(mmse, function(x) {
                 gsub("Correct", 1, x)
            }))
mmse <- data.frame(lapply(mmse, function(x) {
                 gsub("Incorrect", 0, x)
            }))
mmse <- data.frame(lapply(mmse, function(x) {
                 gsub("sc", "bl", x)
            }))
mmse <- data.frame(lapply(mmse, function(x) {
                 gsub("f", "bl", x)
            }))
mmse <- data.frame(lapply(mmse, function(x) {
                 gsub("y1", "m12", x)
            }))
mmse[,subItemScores] = lapply(mmse[,subItemScores], as.character)
mmse[,subItemScores] = lapply(mmse[,subItemScores], as.numeric)

 

collaborativeScoresMMSE= c("MMSE_Orientation", "MMSE_Working_Memory_Registration", "MMSE_Attention_Concentration", "MMSE_Memory_Recall","MMSE_Language")
mmse["MMSE_TIME"] = rowSums(mmse[,c("MMDATE", "MMYEAR", "MMMONTH", "MMDAY", "MMSEASON")])
mmse["MMSE_PLACE"] = rowSums(mmse[,c("MMHOSPIT", "MMFLOOR", "MMCITY", "MMAREA", "MMSTATE")])
mmse["MMSE_Working_Memory_Registration"] = rowSums(mmse[,c("MMBALL", "MMFLAG", "MMTREE")])
mmse["MMSE_Attention_Concentration"] = rowSums(mmse[,c("MMD", "MML", "MMR", "MMO", "MMW")])
mmse["MMSE_Memory_Recall"] = rowSums(mmse[,c("MMBALLDL", "MMFLAGDL", "MMTREEDL")])
mmse["MMSE_Language"] = rowSums(mmse[,c("MMWATCH", "MMPENCIL")])
mmse["MMSE_REPETITION"] =  mmse[,"MMREPEAT"]
mmse["MMSE_COMMANDS"] =rowSums(mmse[,c("MMHAND", "MMFOLD", "MMONFLR", "MMREAD", "MMWRITE", "MMDRAW")])
mmse["MMSE_Language"] = mmse$MMSE_Language +  mmse$MMSE_REPETITION +  mmse$MMSE_COMMANDS
mmse["MMSE_Orientation"] = mmse$MMSE_TIME +  mmse$MMSE_PLACE
```


```{r}
mmse_subscores <- mmse[,c("RID","VISCODE","MMSE_Orientation", "MMSE_Working_Memory_Registration", "MMSE_Attention_Concentration", "MMSE_Memory_Recall","MMSE_Language")]

adni_cog <- merge(adni.features, mmse_subscores, by = c("RID","VISCODE"))
```

###CDR Features###
```{r}
load(paste(getwd(), "/cdr.rdata", sep = ""))
cdr = cdr

cdr <- data.frame(lapply(cdr, function(x) {
                 gsub("sc", "bl", x)
            }))
cdr <- data.frame(lapply(cdr, function(x) {
                 gsub("f", "bl", x)
            }))
cdr <- data.frame(lapply(cdr, function(x) {
                 gsub("y1", "m12", x)
            }))
cdr <- data.frame(lapply(cdr, function(x) {
                 gsub("y2", "m24", x)
            }))
cdr <- data.frame(lapply(cdr, function(x) {
                 gsub("y3", "m36", x)
            }))


cdr_features <- cdr[,c("RID","VISCODE","CDCARE","CDCOMMUN","CDHOME","CDJUDGE","CDMEMORY","CDORIENT")]
adni_cdr <- merge(adni_cog, cdr_features, by = c("RID","VISCODE"))
```

```{r}
load(paste(getwd(), "/adas.rdata", sep = ""))
adasSubSetAdni = adas[,c("RID", "VISCODE", grep("Q[0-9]*SCORE$", colnames(adas), value =TRUE))]
##memory: add Q1SCORE, Q4SCORE, Q7SCORE, Q8SCORE, Q9SCORE
adasSubSetAdni$ADAS_Memory = rowSums(adasSubSetAdni[,grep("Q(1|4|7|8|9)SCORE$", colnames(adasSubSetAdni), value = TRUE)])
##language: add Q2SCORE, Q5SCORE, Q10SCORE, Q11SCORE, Q12SCORE
adasSubSetAdni$ADAS_Language = rowSums(adasSubSetAdni[,grep("Q(2|5|10|11|12)SCORE$", colnames(adasSubSetAdni), value = TRUE)])
##praxis: add Q3SCORE, Q6SCORE 
adasSubSetAdni$ADAS_Praxis = rowSums(adasSubSetAdni[,grep("Q(3|6)SCORE$", colnames(adasSubSetAdni), value = TRUE)])
##concentration:  Q13SCORE
adasSubSetAdni$ADAS_Concentration = adasSubSetAdni[,"Q13SCORE"]
adasVarsADNI = adasSubSetAdni[,c("RID", "VISCODE", "ADAS_Memory", "ADAS_Language", "ADAS_Praxis", "ADAS_Concentration")]
combinedDataADNI <- merge(adni_cdr, adasVarsADNI,by = c("RID", "VISCODE"))
```

###FAQ Features###
```{r}
library(stringr)
faq.features <- faq[,c("RID","VISCODE","FAQFINAN","FAQFORM","FAQSHOP","FAQGAME","FAQBEVG","FAQMEAL","FAQEVENT",
                      "FAQTV","FAQREM","FAQTRAVL","FAQTOTAL")]
faqTests <- faq.features %>% select(starts_with("FAQ"))
faqTests<- colnames(faqTests)
num.extract <- function(string){
  str_extract(string, "\\-*\\d+\\.*\\d*")
  string <- as.numeric(str_extract(string, "\\-*\\d+\\.*\\d*"))
}

faq.features[,c(faqTests)] <- lapply(faq.features[,c(faqTests)],num.extract)
combinedAdniFaq <- merge(combinedDataADNI, faq.features, by = c("RID","VISCODE"))

combinedAdniFaq <- transform(combinedAdniFaq, VISCODE = ifelse(VISCODE != "bl",num.extract(VISCODE), 0))
combinedAdniFaq <- combinedAdniFaq[order(as.integer(combinedAdniFaq$VISCODE)),]
```


```{r}
length(unique(combinedAdniFaq$RID))
```

```{r}
table(combinedAdniFaq$PTETHCAT)
length(combinedAdniFaq$PTETHCAT)
10334/10716
```

```{r}
apply(combinedAdniFaq, 2, function(col)sum(is.na(col))/length(col)*100)
```

```{r}
cleaned.AdniFaq <- na.omit(combinedAdniFaq)
```

```{r}
length(unique(cleaned.AdniFaq$RID))
```

```{r}
print("Number of missing values")
print(sum(is.na(combinedAdniFaq)))
print("Number of rows with missing values")
print(sum(!complete.cases(combinedAdniFaq)))
print("Number of records before cleaning")
print(length(combinedAdniFaq$RID))
print("Number of records after cleaning")
print(length(cleaned.AdniFaq$RID))
```


```{r}
modify.educat <- function(value){
  ifelse(value <= 9, value<-"Low",value<-"High")
}
```


```{r}
baseline.adni <- cleaned.AdniFaq[cleaned.AdniFaq$VISCODE == 0,]
baseline.adni <- baseline.adni[,c("DX","PTGENDER","PTEDUCAT","PTETHCAT","PTRACCAT","PTMARRY","APOE4","AGE")]
#baseline.adni$PTEDUCAT <- sapply(baseline.adni$PTEDUCAT,modify.educat)
```

### Statistics ###
```{r}
ADNI.baseline <- baseline.adni[,c("DX","PTGENDER","PTEDUCAT","PTETHCAT","PTRACCAT","PTMARRY","APOE4","AGE")]

table_summary <- list(
  "Age" = list(
    "Mean (sd)" = ~ mean_sd(AGE)
  ),
  "Gender" = list(
    "Male" = ~ qwraps2::n_perc(PTGENDER == "Male"),
    "Female" = ~ qwraps2::n_perc(PTGENDER == "Female")
  ),
  "Ethnicity" = list(
    "Hisp/Latino" = ~ qwraps2::n_perc(PTETHCAT == "Hisp/Latino"),
    "Not Hisp/Latino" = ~ qwraps2::n_perc(PTETHCAT == "Not Hisp/Latino"),
    "Unknown" = ~ qwraps2::n_perc(PTETHCAT == "Unknown")
  ),
  "Race" = list(
    "White" = ~ qwraps2::n_perc(PTRACCAT == "White"),
    "Black" = ~ qwraps2::n_perc(PTRACCAT == "Black"),
    "Asian" = ~ qwraps2::n_perc(PTRACCAT == "Asian"),
    "More than one" = ~ qwraps2::n_perc(PTRACCAT == "More than one")
  ),
  "Marital Status" = list(
    "Married" = ~ qwraps2::n_perc(PTMARRY == "Married"),
    "Widowed" = ~ qwraps2::n_perc(PTMARRY == "Widowed"),
    "Divorced" = ~ qwraps2::n_perc(PTMARRY == "Divorced"),
    "Never married" = ~ qwraps2::n_perc(PTMARRY == "Never married")
  ),
  "Education (years)" = list(
    "Mean (sd)" = ~ mean_sd(PTEDUCAT)
  ),
  "APOE" = list(
    "Negative" = ~ qwraps2::n_perc(APOE4 == 0),
    "One Allele" = ~ qwraps2::n_perc(APOE4 == 1),
    "Two Alleles" = ~ qwraps2::n_perc(APOE4 == 2)
  )
)
```

### Table Generation ###
```{r, results="asis"}
orig_opt <- options()$qwraps2_markup
options(qwraps2_markup = "markdown")


whole_table <- summary_table(ADNI.baseline,table_summary,by="DX")
print(whole_table, rtitle = "Summary Statistics \1 \1 \1 \1 \1")
```


### Grouping of patients by progression patterns ###
```{r}
library(comprehenr)

aggDiagnosis <- ddply(cleaned.AdniFaq, .(RID), summarize, DX = toString(DX))

onlyCN <-c()
onlyDementia <- c()
convertMCItoDementia <- c()
convertCNtoMCI <- c()
directCNtoDementia <- c()
convertCNtoMCItoDementia <- c()
onlyMCI <- c()
oneCN <- c()
oneMCI <- c()
oneDementia <- c()

for(i in 1:nrow(aggDiagnosis)){
  if(aggDiagnosis[i,2] == "CN"){
    oneCN <- c(oneCN, aggDiagnosis[i,1])
  }
  if(aggDiagnosis[i,2] == "MCI"){
    oneMCI <- c(oneMCI, aggDiagnosis[i,1])
  }
  if(aggDiagnosis[i,2] == "Dementia"){
    oneDementia <- c(oneDementia, aggDiagnosis[i,1])
  }
  if(grepl("^CN((?!MCI|Dementia).)*CN$",aggDiagnosis[i,2],perl=TRUE)){
    onlyCN <- c(onlyCN, aggDiagnosis[i,1])
  }
  if(grepl("^(?!.*MCI)^Dementia.*Dementia$",aggDiagnosis[i,2],perl=TRUE)){
    onlyDementia <- c(onlyDementia, aggDiagnosis[i,1])
  }
  if(grepl("(?!.*CN)^MCI, (MCI, )*(Dementia, )*Dementia$",aggDiagnosis[i,2], perl=TRUE)){
    convertMCItoDementia <- c(convertMCItoDementia, aggDiagnosis[i,1])
  }
  if(grepl("(?!.*Dementia)^CN, (CN, )*(MCI, )*MCI$",aggDiagnosis[i,2],perl=TRUE)){
    convertCNtoMCI <- c(convertCNtoMCI, aggDiagnosis[i,1])
  }
  if(grepl("^(?!.*MCI)^CN, (CN, )*(Dementia, )*Dementia$",aggDiagnosis[i,2], perl=TRUE)){
    directCNtoDementia <- c(directCNtoDementia, aggDiagnosis[i,1])
  }
  if(grepl("^CN, (CN, )*(MCI, )*(Dementia, )*Dementia$",aggDiagnosis[i,2], perl=TRUE)){
    convertCNtoMCItoDementia <- c(convertCNtoMCItoDementia, aggDiagnosis[i,1])
  }
  if(grepl("^MCI((?!CN|Dementia).)*MCI$",aggDiagnosis[i,2],perl=TRUE)){
    onlyMCI <- c(onlyMCI, aggDiagnosis[i,1])
  }
}
selectedIDS <- c(onlyDementia,convertMCItoDementia,directCNtoDementia,convertCNtoMCI,
                convertCNtoMCItoDementia,onlyMCI,onlyCN,oneCN,oneMCI,oneDementia)
#selectedIDS <- c(convertMCItoDementia,
#                 convertCNtoMCItoDementia)
Diagnosis <- c("Only Dementia","Convert MCI to Dementia","Direct CN to Dementia","Convert CN to MCI",
               "Convert CN to MCI to Dementia","Only MCI","Only CN","One MCI","One Dementia")
Count <- c(length(onlyDementia),length(convertMCItoDementia),length(directCNtoDementia),length(convertCNtoMCI),
           length(convertCNtoMCItoDementia),length(onlyMCI),length(onlyCN),length(oneCN),length(oneMCI),
           length(oneDementia))

#Count <- c(length(convertMCItoDementia),
#           length(convertCNtoMCItoDementia))

finalDf <- cleaned.AdniFaq[which(cleaned.AdniFaq$RID %in% selectedIDS),]
```

```{r}
print("Number of records after removing converters")
print(length(finalDf$RID))
table(finalDf$DX)
```


### Filtering last MCI/first Dementia points ###
```{r}
mci <- finalDf[finalDf$DX == "MCI",]
dementia <- finalDf[finalDf$DX == "Dementia",]
```

```{r}
last_mci <- mci %>% group_by(RID) %>% slice(which.max(VISCODE))
first_dementia <- dementia %>% group_by(RID) %>% slice(which.min(VISCODE))
```

```{r}
threshold_df <- rbind(last_mci,first_dementia)
```

```{r}
table(threshold_df$DX)
```

```{r}
adni.split.count <- ddply(threshold_df, .(RID), summarize, DX = toString(DX))
converters <- adni.split.count[adni.split.count$DX == "MCI, Dementia",]
mciToADConverters <- threshold_df[which(threshold_df$RID %in% converters$RID),]
```
### Statistics Table ###
```{r}
ADNI <- mciToADConverters[,c("DX","PTGENDER","PTEDUCAT","PTETHCAT","PTRACCAT","PTMARRY","APOE4","AGE")]

table_summary <- list(
  "Diagnosis" = list(
    "MCI" = ~ qwraps2::n_perc(DX == "MCI"),
    "AD" = ~ qwraps2::n_perc(DX == "Dementia")
  ),
  "Age" = list(
    "Mean (sd)" = ~ mean_sd(AGE)
  ),
  "Gender" = list(
    "Male" = ~ qwraps2::n_perc(PTGENDER == "Male"),
    "Female" = ~ qwraps2::n_perc(PTGENDER == "Female")
  ),
  "Ethnicity" = list(
    "Hisp/Latino" = ~ qwraps2::n_perc(PTETHCAT == "Hisp/Latino"),
    "Not Hisp/Latino" = ~ qwraps2::n_perc(PTETHCAT == "Not Hisp/Latino"),
    "Unknown" = ~ qwraps2::n_perc(PTETHCAT == "Unknown")
  ),
  "Race" = list(
    "White" = ~ qwraps2::n_perc(PTRACCAT == "White"),
    "Black" = ~ qwraps2::n_perc(PTRACCAT == "Black"),
    "Asian" = ~ qwraps2::n_perc(PTRACCAT == "Asian"),
    "More than one" = ~ qwraps2::n_perc(PTRACCAT == "More than one")
  ),
  "Marital Status" = list(
    "Married" = ~ qwraps2::n_perc(PTMARRY == "Married"),
    "Widowed" = ~ qwraps2::n_perc(PTMARRY == "Widowed"),
    "Divorced" = ~ qwraps2::n_perc(PTMARRY == "Divorced"),
    "Never married" = ~ qwraps2::n_perc(PTMARRY == "Never married")
  ),
  "Education (years)" = list(
    "Mean (sd)" = ~ mean_sd(PTEDUCAT)
  ),
  "APOE" = list(
    "Negative" = ~ qwraps2::n_perc(APOE4 == 0),
    "One Allele" = ~ qwraps2::n_perc(APOE4 == 1),
    "Two Alleles" = ~ qwraps2::n_perc(APOE4 == 2)
  )
)
```

```{r, results="asis"}
orig_opt <- options()$qwraps2_markup
options(qwraps2_markup = "markdown")


whole_table <- summary_table(ADNI,table_summary)
print(whole_table, rtitle = "Summary Statistics \1 \1 \1 \1 \1")
```


### Label Encoding ###
```{r}
mciToADConverters <- threshold_df[which(threshold_df$RID %in% converters$RID),]

mciToADConverters$DX <- factor(mciToADConverters$DX,
                               levels = c("Dementia","MCI"),
                               labels = c(0,1))

mciToADConverters$PTGENDER <- factor(mciToADConverters$PTGENDER,
                               levels = c("Female","Male"),
                               labels = c(0,1))
mciToADConverters$PTGENDER <- as.factor(mciToADConverters$PTGENDER)
mciToADConverters$PTETHCAT <- factor(mciToADConverters$PTETHCAT,
                               levels = c("Hisp/Latino","Not Hisp/Latino","Unknown"),
                               labels = c(0,1,2))
mciToADConverters$PTETHCAT <- as.factor(mciToADConverters$PTETHCAT)
mciToADConverters$PTRACCAT <- factor(mciToADConverters$PTRACCAT,
                               levels = c("Asian","Black","More than one","White"),
                               labels = c(0,1,2,3))
mciToADConverters$PTRACCAT <- as.factor(mciToADConverters$PTRACCAT)
mciToADConverters$PTMARRY <- factor(mciToADConverters$PTMARRY,
                               levels = c("Divorced","Married","Never married","Widowed"),
                               labels = c(0,1,2,3))
mciToADConverters$PTMARRY <- as.factor(mciToADConverters$PTMARRY)
```


### Shuffling ###
```{r}
set.seed(1)
rows <- sample(nrow(mciToADConverters))
shuffleTest <- mciToADConverters[rows,]
write_csv(shuffleTest,"ADNI_shuffle_08_10_2020.csv")
```


analyze.significance = function(vars, mydat, classes, confounders=c(), out="~"){
  # 1. choose "optimal" confounder model
  #m0 = do.call("glm", list(as.formula(paste(classes, "~ .")), data=na.omit(mydat[, c(classes, confounders)])))
  m0Form<- as.formula(paste0(paste("classes", paste(confounders, collapse=" + "), sep=" ~ "), " +(1|RID)"))
  m0 =  glmer(m0Form, data = mydat, family=binomial(link="logit"), control=glmerControl(optimizer="bobyqa"))
  print(m0)
  #print(summary(m0)$coefficients)
  #p0 <- summary(m0)$coefficients[,4]
  conf = names(attr(terms(m0), "dataClass"))
  cat("optimal confounder set:", conf, "\n")
  res = data.frame()
  # 2. test additional influence of variables of interest
  for(v in vars){
    cat("variable = ", v, "\n")
    group <- c(v,conf)
    mForm <- as.formula(paste0(paste("classes", paste(group, collapse=" + "), sep=" ~ "), " +(1|RID)"))
    m =  glmer(mForm, data = mydat, family=binomial(link="logit"), control=glmerControl(optimizer="bobyqa"))
    print(m)
    #m  = do.call("glm", list(as.formula(paste(classes, "~ .")), data=na.omit(mydat[, c(classes, group)])))
    print(summary(m)$coefficients)
    lr <- lrtest(m0,m)
    print(lr)
    p <- summary(m)$coefficients[,4][v]
    A = Anova(m, type="III")[1,3]
    if(A < 0.05){
      res = rbind(res, data.frame(variable=v, confounders = paste(conf, sep="", collapse=", "), p.value=p, p.value.global = A, stringsAsFactors = FALSE))
    }else{
      res = rbind(res, data.frame(variable=v, confounders=NA, p.value=NA, p.value.global=A))
  }}
  res$FDR = p.adjust(res$p.value, method="fdr") 
  res$FDR.round = round(res$FDR, 2)
  res
}

### GLMM and Liklihood Ratio Tests ###
```{r}
analyze.significance = function(vars,mydat,classes,confounders=c(), out="~"){
  #base model
  m0 <- glmer(paste0(paste(classes,paste(confounders,collapse=" + ")
                           ,sep=" ~ "),"+(1|RID)"),data=mydat,family=binomial(link="logit"),control=glmerControl(optimizer="bobyqa"))
  #print(summary(m0)$coefficients)[,4]
  for(v in vars){
    #alternative model
    group <- c(v,confounders) 
    m <- glmer(paste0(paste(classes,paste(group,collapse=" + ")
                           ,sep=" ~ "),"+(1|RID)"),data=mydat,family=binomial(link="logit"),control=glmerControl(optimizer="bobyqa"))
    #anova test
    print(summary(m)$coefficients)
    #print("Anova")
    anova_test = anova(m0, m)
    #print(anova_test)
    print("LR Test")
    lr <- lrtest(m0,m)
    print(lr)
    
    
library(lmtest)
library(lme4)
#shuffleTest$DX = as.numeric(shuffleTest$DX)
test = shuffleTest

confounders = c("PTGENDER","PTEDUCAT","PTETHCAT","PTRACCAT","PTMARRY","APOE4","AGE")
measures = c("FAQFINAN","FAQFORM","FAQSHOP","FAQGAME","FAQBEVG","FAQMEAL","FAQEVENT",
             "FAQTV","FAQREM","FAQTRAVL")
analyze.significance(measures,test,"DX",confounders,out="~")
```


library(lmtest)
library(lme4)
shuffleTest$DX = as.numeric(shuffleTest$DX)
test = shuffleTest
confounders = c(test$PTGENDER,test$PTEDUCAT,test$PTETHCAT,test$PTRACCAT,test$PTMARRY,test$APOE4,test$AGE)
measures = c("FAQFINAN","FAQFORM","FAQSHOP","FAQGAME","FAQBEVG","FAQMEAL","FAQEVENT",
             "FAQTV","FAQREM","FAQTRAVL")
analyze.significance(measures,test,"DX",confounders = c("PTGENDER","PTEDUCAT",
                                                        "PTETHCAT","PTRACCAT","PTMARRY","APOE4","AGE"), out="~")

```{r}
write_csv(mciToADConverters,"ADNI_final_04_09_2020.csv")
```