---
title: "dataPreparation"
output:
  pdf_document: default
  html_document: default
---
```{r}
rm(list=ls())
library(plyr)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set('dataPreparationADNI.Rmd', encoding = 'UTF-8')
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = getwd())
```

##adni data##
```{r}
load(paste(getwd(), "/ADNIMERGE_Aug2020/data/adnimerge.rdata", sep = ""))
adnimerge$long.age <- round((adnimerge[,"AGE"] + adnimerge[,"Years.bl"]), digits =1)
```

##Normalizing the brain volumes by ICV
```{r}
brainVolumesNotICV <- c("Ventricles", "Hippocampus", "WholeBrain", "Entorhinal", "Fusiform", "MidTemp")
funcDividebyICV = function(df){
  for(cols in brainVolumesNotICV){
    df[,cols] = df[,cols]/df[,"ICV"]
  }
  return(df)
}

adnimerge = funcDividebyICV(adnimerge)
```

##number of patients##
```{r}
length(unique(adnimerge$RID))
```

##diagnostic distribution at baseline##
```{r}
print("All patients in total are:")
adnimergeAtBl = subset(adnimerge, adnimerge$Month == 0)
length(unique(adnimergeAtBl$RID))
table(adnimergeAtBl$DX.bl)
```

```{r}
##Choose which type of measures you want to do pattern analysis for
# func.typeof.data <- function()
# {
#   user_input <- "Do you want to do trajectory predictions  for :
#   1. MCIandDementia
#   2. DeNovo
#   3. Control
#   4. MCIandDementiaFAQ
#   "
#   type_of_data <- readline(prompt = user_input)
#   return(type_of_data)
# }
# type_of_data <- func.typeof.data()
```

```{r}
##extract the IDs that are there in Salah's data
#ADNI_IDs <- read_csv("~/Documents/Documents_IT/ADNIVAMBN/VAMBNForADNI/ADNI_IDs.csv")
#final_data_classifier = read_csv("ADNI_shuffle_08_10_2020.csv")
finalDf = adnimerge
```

```{r}
##remove the patients that revert from Dementia to MCI or CN
aggDiagnosis <- ddply(finalDf, .(RID), summarise, DX = toString(DX))

func.revert <- function(aggDf){
    patientidConv <- c()
for(i in 1:nrow(aggDf)){
  if(grepl("^Dementia(.*MCI|.*\\,\\sMCI|\\,\\sMCI|.*\\,\\sMCI$|.*CN|.*\\,\\sCN|\\,\\sCN|.*\\,\\sCN$)", aggDf[i,2])){
      patientidConv <- c(patientidConv, aggDf[i,1])
  }
   # if(grepl("^MCI(.*CN|\\,\\sCN|.*\\,\\sCN|.*\\,\\sMCI$)", aggDf[i,2])){
   #    patientidConv <- c(patientidConv, aggDf[i,1])
   # }
     if(grepl("^CN.*MCI.*Dementia(.*CN|\\,\\sCN|.*\\,\\sCN|.*\\,\\sMCI$)", aggDf[i,2])){
      patientidConv <- c(patientidConv, aggDf[i,1])
     }
  if(grepl("^MCI.*Dementia(.*CN|\\,\\sCN|.*\\,\\sCN|.*\\,\\sMCI$)", aggDf[i,2])){
      patientidConv <- c(patientidConv, aggDf[i,1])
     }
  
}
  print("Patients reverting")
  print(patientidConv)
  return(patientidConv)
}


# if(type_of_data %in%  c("MCIandDementia","MCIandDementiaFAQ")){
#   patientid <- c()
#   for(i in 1:nrow(aggDiagnosis)){
#     if(grepl("Dementia", aggDiagnosis[i,2])){
#        patientid <- c(patientid, aggDiagnosis[i,1])
#     }
#   }
#   atleastOneADdf <- adnimerge[which(adnimerge$RID %in% patientid),]
#   aggDiagnosisAD <- ddply(atleastOneADdf, .(RID), summarise, DX = toString(DX))
#   patientidMCI <- c()
#   for(i in 1:nrow(aggDiagnosisAD)){
#     if(grepl("MCI", aggDiagnosisAD[i,2])){
#         patientidMCI <- c(patientidMCI, aggDiagnosisAD[i,1])
#     }
#   }
#   atleastOneMCIdf <- atleastOneADdf[which(atleastOneADdf$RID %in% patientidMCI),]
#   aggDiagnosisMCI <- ddply(atleastOneMCIdf, .(RID), summarise, DX = toString(DX))
#   patientidConv <- func.revert(aggDiagnosisMCI)
#   ##remove these ids from the df as they are recoverting to MCI or CTL from AD and CTL from MCI##
#   finalDf <- atleastOneMCIdf[which(!atleastOneMCIdf$RID %in% patientidConv),]
# }
# 
# if(type_of_data == "DeNovo"){
#   ##patients with subjects having Dementia diagnosis throughout
#   onlyDementia <- c()
#   for(i in 1:nrow(aggDiagnosis)){
#     if(grepl("^(?!.*MCI)^Dementia.*Dementia$",aggDiagnosis[i,2],perl=TRUE)){
#     onlyDementia <- c(onlyDementia, aggDiagnosis[i,1])
#       }
#   }
#   deNovoDementia <- adnimerge[which(adnimerge$RID %in% onlyDementia),]
#   aggDiagnosisDementia <- ddply(deNovoDementia, .(RID), summarize, DX = toString(DX))
#   patientidConv <- func.revert(aggDiagnosisDementia)
#   finalDf <- deNovoDementia[which(!deNovoDementia$RID %in% patientidConv),]
# }
# 
# if(type_of_data == "Control"){
#   ##control subjects##
#   onlyCN<- c()
#   for(i in 1:nrow(aggDiagnosis)){
#     if(grepl("^(?!.*MCI|Dementia)^CN.*CN$",aggDiagnosis[i,2],perl=TRUE)){
#       onlyCN <- c(onlyCN, aggDiagnosis[i,1])
#     }
#   }
#   control <- adnimerge[which(adnimerge$RID %in% onlyCN),]
#   aggDiagnosisControl <- ddply(control, .(RID), summarize, DX = toString(DX))
#   patientidConv <- func.revert(aggDiagnosisControl)
#   finalDf <- control[which(!control$RID %in% patientidConv),]
# }
patientidConv <- func.revert(aggDiagnosis)
finalDf <- finalDf[which(!finalDf$RID %in% patientidConv),]
length(unique(finalDf$RID))
```

##patients converting from NL to MCI to Dementia
```{r}
ts <- ddply(finalDf, .(RID), summarise, DX = toString(DX))
nlSubjects <- subset(ts, grepl("CN", ts$DX, perl = TRUE))
length(unique(nlSubjects))
```

##extract only cognitive scores from adnimerge##
```{r}
mri <- c("Ventricles", "Hippocampus", "WholeBrain", "Entorhinal", "Fusiform", "MidTemp", "ICV")
#ravlt <- grep("^RAVLT.*[^bl]$", colnames(finalDf), value = TRUE)
#ecog <- grep("^Ecog.*[^bl]$",  colnames(adnimerge), value = TRUE)
#cognition <- c("MMSE", "CDRSB", "ADAS13", "ADAS11", "MOCA", "LDELTOTAL", "DIGITSCOR", "TRABSCOR", "mPACCdigit", "mPACCtrailsB")
csf <- c("ABETA", "TAU", "PTAU")
imagingPET <- c("PIB","FDG","AV45")
finalDf <- finalDf[,c("RID", "PTID", "AGE", "PTGENDER",  "PTEDUCAT", "PTETHCAT", "PTRACCAT", "PTMARRY", "APOE4", "VISCODE", "ORIGPROT", "COLPROT", "DX",imagingPET, mri,csf)]
```

##add polygenic risk scores from a table##
```{r}
load(paste(getwd(), "/ADNIMERGE_Aug2020/data/desikanlab.rdata", sep = ""))
finalDf <- merge(finalDf, desikanlab[,c("RID", "PHS")])
```

##merge adas individual scores to the table##
```{r}
##for ADNI
##subset adas table for subscores Q scores:
# load(paste(getwd(), "/ADNIMERGE_Aug2020/data/adas.rdata", sep = ""))
# adas$VISCODE[adas$RID == "1052" & adas$USERDATE == "2019-04-02"] <- "m156" #This is hard coded as it's a mistake in original dataset.
# adas$VISCODE[adas$RID == "4175" & adas$USERDATE == "2019-04-24"] <- "m96"
# adas$VISCODE[adas$RID == "4926" & adas$USERDATE == "2019-03-20"] <- "m81"
# adasSubSetAdni = adas[,c("RID", "VISCODE", "ORIGPROT", "COLPROT", grep("Q[0-9]*SCORE$", colnames(adas), value =TRUE))]
# ##memory: add Q1SCORE, Q4SCORE, Q7SCORE, Q8SCORE, Q9SCORE
# adasSubSetAdni$ADAS_Memory = rowSums(adasSubSetAdni[,grep("Q(1|4|7|8|9)SCORE$", colnames(adasSubSetAdni), value = TRUE)])
# ##language: add Q2SCORE, Q5SCORE, Q10SCORE, Q11SCORE, Q12SCORE
# adasSubSetAdni$ADAS_Language = rowSums(adasSubSetAdni[,grep("Q(2|5|10|11|12)SCORE$", colnames(adasSubSetAdni), value = TRUE)])
# ##praxis: add Q3SCORE, Q6SCORE 
# adasSubSetAdni$ADAS_Praxis = rowSums(adasSubSetAdni[,grep("Q(3|6)SCORE$", colnames(adasSubSetAdni), value = TRUE)])
# ##concentration:  Q13SCORE
# adasSubSetAdni$ADAS_Concentration = adasSubSetAdni[,"Q13SCORE"]
# adasVarsADNI = adasSubSetAdni[,c("RID", "VISCODE", "ORIGPROT", "COLPROT", "ADAS_Memory", "ADAS_Language", "ADAS_Praxis", "ADAS_Concentration")]
# combinedDataADNI <- merge(adasVarsADNI, finalDf,by = c("RID", "VISCODE", "ORIGPROT", "COLPROT"))
# length(unique(combinedDataADNI$RID))
combinedDataADNI = finalDf
```

##merge mmse individual scores to the table##
```{r}
##for ADNI
##subset adas table for subscores Q scores:
load(paste(getwd(), "/ADNIMERGE_Aug2020/data/mmse.rdata", sep = ""))
mmse = mmse
#mmse[mmse=="Incorrect"]<-0
subItemScores = c("MMDATE", "MMYEAR", "MMMONTH", "MMDAY", "MMSEASON", "MMHOSPIT", "MMFLOOR", "MMCITY", "MMAREA", "MMSTATE", "MMBALL", "MMFLAG", "MMTREE", "MMD", "MML", "MMR", "MMO", "MMW", "MMBALLDL", "MMFLAGDL", "MMTREEDL", "MMWATCH", "MMPENCIL", "MMREPEAT", "MMHAND", "MMFOLD", "MMONFLR", "MMREAD", "MMWRITE", "MMDRAW")

mmse <- data.frame(lapply(mmse, function(x) {
                 gsub("0 - incorrect", 0, x)
            }))
mmse <- data.frame(lapply(mmse, function(x) {
                 gsub("0 - Incorrect",0, x)
            }))
mmse <- data.frame(lapply(mmse, function(x) {
                 gsub("1 - Correct", 1, x)
            }))
mmse <- data.frame(lapply(mmse, function(x) {
                 gsub("1 - correct", 1, x)
            }))
mmse <- data.frame(lapply(mmse, function(x) {
                 gsub("Correct", 1, x)
            }))
mmse <- data.frame(lapply(mmse, function(x) {
                 gsub("Incorrect", 0, x)
            }))
mmse <- data.frame(lapply(mmse, function(x) {
                 gsub("sc", "bl", x)
            }))
mmse <- data.frame(lapply(mmse, function(x) {
                 gsub("f", "bl", x)
            }))
mmse <- data.frame(lapply(mmse, function(x) {
                 gsub("y1", "m12", x)
            }))
mmse[,subItemScores] = lapply(mmse[,subItemScores], as.character)
mmse[,subItemScores] = lapply(mmse[,subItemScores], as.numeric)

collaborativeScoresMMSE= c("MMSE_Orientation", "MMSE_Working_Memory_Registration", "MMSE_Attention_Concentration", "MMSE_Memory_Recall","MMSE_Language")
mmse["MMSE_TIME"] = rowSums(mmse[,c("MMDATE", "MMYEAR", "MMMONTH", "MMDAY", "MMSEASON")])
mmse["MMSE_PLACE"] = rowSums(mmse[,c("MMHOSPIT", "MMFLOOR", "MMCITY", "MMAREA", "MMSTATE")])
mmse["MMSE_Working_Memory_Registration"] = rowSums(mmse[,c("MMBALL", "MMFLAG", "MMTREE")])
mmse["MMSE_Attention_Concentration"] = rowSums(mmse[,c("MMD", "MML", "MMR", "MMO", "MMW")])
mmse["MMSE_Memory_Recall"] = rowSums(mmse[,c("MMBALLDL", "MMFLAGDL", "MMTREEDL")])
mmse["MMSE_Language"] = rowSums(mmse[,c("MMWATCH", "MMPENCIL")])
mmse["MMSE_REPETITION"] =  mmse[,"MMREPEAT"]
mmse["MMSE_COMMANDS"] =rowSums(mmse[,c("MMHAND", "MMFOLD", "MMONFLR", "MMREAD", "MMWRITE", "MMDRAW")])
mmse["MMSE_Language"] = mmse$MMSE_Language +  mmse$MMSE_REPETITION +  mmse$MMSE_COMMANDS
mmse["MMSE_Orientation"] = mmse$MMSE_TIME +  mmse$MMSE_PLACE
combinedDataADNImmse <- merge(mmse[c("RID", "VISCODE", "ORIGPROT", "COLPROT", collaborativeScoresMMSE)], combinedDataADNI,by = c("RID", "VISCODE", "ORIGPROT", "COLPROT"))
length(unique(combinedDataADNImmse$RID))
```

##cdrsb individual scores to the table##
```{r}
##for CDRSB
##subset faq table for subscores Q scores:
# load(paste(getwd(), "/ADNIMERGE_Aug2020/data/cdr.rdata", sep = ""))
# cdrsbRelevant <- cdr[,c("RID","VISCODE","ORIGPROT","COLPROT","CDCARE", "CDCOMMUN", "CDHOME", "CDJUDGE",
#                         "CDMEMORY", "CDORIENT")]
# combinedAdniCDRSB <- merge(combinedDataADNImmse, cdrsbRelevant, by = c("RID","VISCODE", "ORIGPROT", "COLPROT"))
# length(unique(combinedAdniCDRSB$RID))
```



##merge faq individual scores to the table##
```{r}
#library(tidyverse)
library(dplyr)
##for ADNI
##subset faq table for subscores Q scores:
load(paste(getwd(), "/ADNIMERGE_Aug2020/data/faq.rdata", sep = ""))
faq$VISCODE[faq$RID == "1052" & faq$USERDATE == "2019-03-29"] <- "m156"
faq$VISCODE[faq$RID == "4175" & faq$USERDATE == "2019-03-18"] <- "m96"
faq$VISCODE[faq$RID == "5075" & faq$USERDATE == "2018-12-05"] <- "m72"
faq$VISCODE[faq$RID == "4926" & faq$USERDATE == "2019-03-21"] <- "m81"
#faq[2308,5] <- "m106"
print("Number of Patients:")
length(unique(faq$RID))
library(dplyr)
library(stringr)
faqRelevant <- faq[,c("RID","VISCODE","ORIGPROT","COLPROT","FAQFINAN","FAQFORM","FAQSHOP","FAQGAME","FAQBEVG","FAQMEAL","FAQEVENT","FAQTV","FAQREM","FAQTRAVL","FAQTOTAL")]
faqTests <- faqRelevant %>%  dplyr::select(starts_with("FAQ"))
faqTests<- colnames(faqTests)
num.extract <- function(string){
  str_extract(string, "\\-*\\d+\\.*\\d*")
  string <- as.numeric(str_extract(string, "\\-*\\d+\\.*\\d*"))
}
faqRelevant[,c(faqTests)] <- lapply(faqRelevant[,c(faqTests)],num.extract)
combinedAdniFaq <- merge(combinedDataADNImmse, faqRelevant, by = c("RID","VISCODE", "ORIGPROT", "COLPROT"))
length(unique(combinedAdniFaq$RID))
```


##delete the patients with just one time point##
```{r}
func.del.patients.single.time.point <- function(data){
  library(plyr)
  aggVisit <- ddply(data, .(RID), summarise, VISCODE = toString(VISCODE))
  print(aggVisit)
  data <- subset(data,!data$RID 
                 %in%aggVisit[which(count.fields(textConnection((aggVisit[,"VISCODE"])), sep = ",")== 1), 1])
  #print(nrow(data))
  return(data)
}
combinedAdniFaq <- func.del.patients.single.time.point(combinedAdniFaq)
length(unique(combinedAdniFaq$RID))
```


##converting the long format to wide format#
```{r}
##extracting the columns that need to be converted to wider format##
mmseFeatures = grep("^MMSE", colnames(combinedAdniFaq), value = TRUE)
faqFeatures = grep("^FAQ", colnames(combinedAdniFaq), value = TRUE)
#adasFeatures = grep("^ADAS", colnames(combinedAdniFaq), value = TRUE)
mostRelevantCols <- c(mri, imagingPET, csf, mmseFeatures, faqFeatures)
demogs <- grep("^PT[^ID|AU].*", colnames(combinedAdniFaq), value = TRUE)
demogsAndOthers <- c(demogs, "RID", "APOE4", "ORIGPROT", "COLPROT", "PHS", "AGE")
# CSF biomarkers values are represented as characters due to presence of "<" or ">" symbols
#convert them to numeric
print("converting to Numeric")
combinedAdniFaq$ABETA <- gsub('(<|>)','',combinedAdniFaq$ABETA)
combinedAdniFaq$PTAU <- gsub('(<|>)','',combinedAdniFaq$PTAU)
combinedAdniFaq$TAU <- gsub('(<|>)','',combinedAdniFaq$TAU)

combinedAdniFaq$ABETA <- as.numeric(combinedAdniFaq$ABETA)
combinedAdniFaq$PTAU <- as.numeric(combinedAdniFaq$PTAU)
combinedAdniFaq$TAU <- as.numeric(combinedAdniFaq$TAU)

##extracting the columns that need to be converted to wider format##
selectedCols <- setdiff(c(colnames(combinedAdniFaq)), demogsAndOthers)
convDenovoSelectedDf <- combinedAdniFaq[,selectedCols]
combinedAdniFaqWide = reshape(convDenovoSelectedDf,idvar='PTID',timevar="VISCODE",dir='w')

#convDenovoSelectedDf <- combinedAdniFaq[,demogsAndOthers]

#save.image("~/Documents/ADNIVAMBN/VAMBNForADNI/ADNIVAMBN.RData")
```

##add snp features and baseline brain volumes##
```{r}
library(gdata)
library(data.table)
library(readxl)
##add snp columns##
#allSNPsAndBurdenScore<- read_csv("~/Documents/Documents_IT/snpWork/allSNPsAndBurdenScores.csv")
#allSNPsAndBurdenScore$X1 = NULL
#names(allSNPsAndBurdenScore)[names(allSNPsAndBurdenScore) == "SUBJID"] = "PTID"
#allSNPsAndBurdenScore$PTID = sub("[0-9]+[\\_]", "", allSNPsAndBurdenScore$PTID)
# ##add brain regions
#EMC_ADNI_FS60_Phenotypes_Desikan_20180219<-read.xlsx("/Users/msood/Documents/Documents – IT-Admin’s MacBook Pro/BayesianNetworkAD/henryImagingData/EMC_ADNI_FS60_Phenotypes_Desikan_20180219.xlsx")
#volumeRegion <- as.data.frame(EMC_ADNI_FS60_Phenotypes_Desikan_20180219)
#brainRegion <- c(colnames(volumeRegion)[grep("Left|Right", colnames(volumeRegion))])
#names(volumeRegion)[1] <- "PTID"
#volumeRegion <- volumeRegion[,c("PTID",brainRegion)]
#volumeRegion <- setDT(volumeRegion, keep.rownames = TRUE)[]
#volumeRegion$rn <- NULL
combinedAdniWide = combinedAdniFaqWide
demogsAndOthersDf <- combinedAdniFaq[,c("PTID",demogsAndOthers)]
#combinedAdniWide <- merge(combinedAdniWide, allSNPsAndBurdenScore, all.x= TRUE)

##1.csf data frame##
csfDf <- combinedAdniWide[,grep("PTID|ABETA|TAU|PTAU", colnames(combinedAdniWide), value = TRUE)]

##Volumetric data frame##
volumeDf <- combinedAdniWide[,grep("PTID|Entorhinal|Fusiform|MidTemp|Hippocampus|Ventricles|WholeBrain|ICV",
                                      colnames(combinedAdniWide), value = TRUE)]


##Cognitive test data frame##
#cogTestFaqDf <- combinedAdniWide[,grep("^CDRSB|^ADASQ4|^MMSE\\.|^RAVLT|^MOCA|^LDELTOTAL|^DIGITSCOR|^TRABSCOR|^mPACCdigit|^mPACCtrailsB|^FAQ\\.|ADAS[^0-9|^Q4]|EcogPt|FAQTOTAL", colnames(combinedAdniWide), value = TRUE)]

mmseDf <- combinedAdniWide[,grep("PTID|MMSE\\_", colnames(combinedAdniWide), value = TRUE)]

faqDf <- combinedAdniWide[,grep("PTID|FAQ[A-Z].*\\.", colnames(combinedAdniWide), value = TRUE)]
faqDf <- faqDf %>% dplyr::select(-contains("FAQTOTAL"))

imagingPETDf <- as.data.frame(combinedAdniWide[,grep("PTID|FDG|PET|AV45", colnames(combinedAdniWide), value = TRUE)])

##Diagnostic data frame##
dxDf <- combinedAdniWide[,grep("PTID|DX", colnames(combinedAdniWide), value = TRUE)]


# # Remove columns with more than 50% missing value
csfDf = csfDf[ , -which(colMeans(is.na(csfDf)) > 0.50)]
volumeDf = volumeDf[ , -which(colMeans(is.na(volumeDf)) > 0.50)]
#cogTestFaqDf = cogTestFaqDf[ , -which(colMeans(is.na(cogTestFaqDf)) > 0.50)]
mmseDf = mmseDf[ , -which(colMeans(is.na(mmseDf)) > 0.50)]
faqDf = faqDf[ , -which(colMeans(is.na(faqDf)) > 0.50)]
imagingPETDf = as.data.frame(imagingPETDf[ , -which(colMeans(is.na(imagingPETDf)) > 0.50)])
#names(imagingPETDf)[1] = "FDG.bl"
dxDf = dxDf[ , -which(colMeans(is.na(dxDf)) > 0.50)]

#cortical regions
#corticalDesikanDf <- combinedAdniWide[,brainRegion]

#allSnpsDf = combinedAdniWide[,grep("^rs[0-9]*", colnames(combinedAdniWide), value = TRUE)]
#allSnpsDf = as.data.frame(lapply(allSnpsDf[, grep("^rs[0-9]*", colnames(allSnpsDf), value = TRUE)], factor))
#allSnpsDf = allSnpsDf[ , which(colMeans(is.na(allSnpsDf)) <= 0.5)]
#pathways
#pathwayCols = setdiff(colnames(allSNPsAndBurdenScore), grep("rs", colnames(allSNPsAndBurdenScore), value = TRUE))
#pathwaysDf = combinedAdniWide[,setdiff(pathwayCols, "PTID")]
#pathwaysDf = pathwaysDf[ , -which(colMeans(is.na(pathwaysDf)) > 0.5)]
#relPathwayDf = as.data.frame(colnames(pathwaysDf))
#write.csv(relPathwayDf, "/Users/msood/Documents/Documents_IT/snpWork/relevantPathwaysAD.csv")
```



```{r}
##Add bl extension to the demogsAndOthers features##
##Combine all the data frames##
#allFeatures <- cbind.data.frame(demogsAndOthersDf, csfDf, volumeDf, cogTestDf, fdgDf, dxDf, snpDf, pathDf, corticalDesikanDf)
demogsAndOthersDf$ORIGPROT = NULL
demogsAndOthersDf$COLPROT = NULL
demogsAndOthersDf = unique(demogsAndOthersDf)
combinedAdniWide <- merge(combinedAdniWide, demogsAndOthersDf, all.x = TRUE)
demogsAndOthers <- setdiff(demogsAndOthers, c("VISCDE","PTID", "ORIGPROT", "COLPROT", "RID"))
demogsAndOthersDf <- combinedAdniWide[,c("PTID", demogsAndOthers)]
colnames(demogsAndOthersDf) <- paste(colnames(demogsAndOthersDf) , ".bl", sep = "")
names(demogsAndOthersDf)[1] <- "PTID"
# Add group name to columns
icol <- which(names(csfDf) == "PTID")
colnames(csfDf)[-icol] <- paste("csf",colnames(csfDf)[-icol],  sep = "_")
icol <- which(names(volumeDf) == "PTID")
colnames(volumeDf)[-icol] <- paste("brain",colnames(volumeDf)[-icol],  sep = "_")
#colnames(cogTestFaqDf) = paste0("Cog_",colnames(cogTestFaqDf))
#colnames(corticalDesikanDf) = paste0("cortical_",colnames(corticalDesikanDf), ".bl")
#colnames(allSnpsDf) = paste0("snp_",colnames(allSnpsDf), ".bl")
#colnames(pathwaysDf) = paste0("path_",colnames(pathwaysDf), ".bl")


#colnames(mmseDf) = paste0("MMSE_",colnames(adasDf))
#colnames(faqDf) = paste0("FAQ_",colnames(faqDf))
```


```{r}
# Auxiliary variables keep track of visit-wise and group-wise patient dropout. 
# Measurements of features are marked by value missing not at random (MNAR).
# MNAR  results from a systematic absence of subject data for a measurement type (feature group). 
# get_aux_all_groups = function(cohortdata){
#   mysample = cohortdata
#   timepoint = str_extract(colnames(mysample), "m[0-9][0-9]")[1]
#   print(timepoint)
#   csf = select(mysample,grep( "csf",colnames(mysample),value=TRUE)) 
#   #print(csf)
#   volumes = select(mysample,grep( "brain",colnames(mysample),value=TRUE)) 
#   #print(volumes)
#   cogTest = select(mysample,grep( "Cog",colnames(mysample),value=TRUE)) 
#   adas = select(mysample,grep( "ADAS",colnames(mysample),value=TRUE)) 
#   #print(cogTest)
#   snp = select(mysample,grep( "snp",colnames(mysample),value=TRUE)) 
#   path = select(mysample,grep( "path",colnames(mysample),value=TRUE)) 
#   cort = select(mysample,grep( "cortical",colnames(mysample),value=TRUE)) 
#   output_aux = function(mysubsample){
#     
#     #return_df = data.frame()
#     #groupname = deparse(substitute(a))
#     #print("nsbsn")
#     if(dim(mysubsample)[2] != 0 ){
#       print("jjdjd")
#       if(dim(mysubsample)[2]== 1){
#         print("memem")
#         mysubsample = mysubsample
#       }else{
#         # Add a new column for AUX
#         new = "new"
#         in.loop = mysubsample
#         #print("jsksks")
#         #print(in.loop)
#         mysubsample[new] <- 0
#         
#         # Get rownames where all value is NA 
#         mysubsample_NA = which(apply(in.loop, 1, function(x) all(is.na(x))))
#         mysubsample_pat = names(mysubsample_NA)
#         #print(mysubsample_pat)
#         print("meemansa")
#         #print(mysubsample_pat)
#         if(length(mysubsample_pat) !=0 ){
#           print("wewewew")
#           mysubsample[which(rownames(mysubsample) %in% mysubsample_pat ),]$new <- 1
#           #print(mysubsample)
#           # Annoate aux column with group name and visit number 
#           groupName = sub("_.*$", "", colnames(mysubsample)[1])
#           print("meem1")
#           print(groupName)
#           new2 = paste(groupName,"aux",timepoint, sep = "_")
#           colnames(mysubsample)[which(names(mysubsample) == "new")] <- new2
#           print(paste0("Aux available for ", groupName, "at" ,timepoint ))
#         }else{ 
#           mysubsample$new  = NULL
#           groupName = sub("_.*$", "", colnames(mysubsample)[1])
#           print("meem2")
#           print(groupName)
#           print(paste0("Aux unavailable for ", groupName, "at" ,timepoint ))}
#       }
#       
#     }else{  print(paste0("Missing group at",timepoint ))}
#     
#     return(mysubsample)
#   }
#   
#   csf_aux = output_aux(mysubsample = csf)
#   #print("nac")
#   #print(csf_aux)
#   volume_aux = output_aux(mysubsample = volumes)
#   cogTest_aux = output_aux(mysubsample = cogTest)
#   adas_aux = output_aux(mysubsample = adas)
#   snp_aux = output_aux(mysubsample = snp)
#   path_aux = output_aux(mysubsample = path)
#   cort_aux = output_aux(mysubsample = cort)
#   #diagnostics_aux = output_aux(mysubsample = diagnostics)
#   
#   outputdf <- data.frame(matrix("removelater", ncol = 1, nrow = nrow(mysample)))
#   names(outputdf)[1]<- "toremove"
#   if(dim(csf)[2] != 0 ){
#     outputdf = as.data.frame(cbind(outputdf , csf_aux))
#     #print(outputdf)
#   } else{
#     #print(paste0("CSF data unavailable for visit", timepoint))
#   }
#   if(dim(volumes)[2] != 0 ){
#     outputdf = as.data.frame(cbind(outputdf , volume_aux))
#   } else{
#     #print(paste0("Volumetric data unavailable for visit", timepoint))
#   }
#   if(dim(cogTest)[2] != 0 ){
#     outputdf = as.data.frame(cbind(outputdf , cogTest_aux))
#   } else{
#    # print(paste0("Cognition test unavailable for visit", timepoint))
#   }
#   if(dim(adas)[2] != 0 ){
#     outputdf = as.data.frame(cbind(outputdf , adas_aux))
#   } else{
#    # print(paste0("Cognition test unavailable for visit", timepoint))
#   }
#   if(dim(snp)[2] != 0 ){
#     outputdf = as.data.frame(cbind(outputdf , snp_aux))
#   } else{
#    # print(paste0("Snp unavailable for visit", timepoint))
#   }
#   if(dim(path)[2] != 0 ){
#     outputdf = as.data.frame(cbind(outputdf , path_aux))
#   } else{
#    # print(paste0("Snp unavailable for visit", timepoint))
#   }
#   if(dim(cort)[2] != 0 ){
#     outputdf = as.data.frame(cbind(outputdf , cort_aux))
#   } else{
#     #print(paste0("Cortical brain region unavailable for visit", timepoint))
#   }
#   outputdf$toremove = NULL
#   return(outputdf)
# }
```


```{r}
#allData = merge(csfDf, volumeDf,mmseDf, faqDf, imagingPETDf, by = "PTID", all = TRUE)
allData    =  Reduce(function(x, y) merge(x, y, all=TRUE), list(csfDf, volumeDf,mmseDf, faqDf, imagingPETDf))
#allData = lapply(allData, as.integer)

allData <- merge(allData, dxDf)
#allData <- merge(allData, demogsAndOthersDf)
visitData = list("visitbl" = allData[, grep(".bl", colnames(allData), value = TRUE)],
                 "visit1" = allData[, grep(".m06", colnames(allData), value = TRUE)],
                 "visit2" = allData[, grep(".m12", colnames(allData), value = TRUE)],
                 "visit3" = allData[, grep(".m24", colnames(allData), value = TRUE)],
                 "visit4" = allData[, grep(".m36", colnames(allData), value = TRUE)]
                )
                 # "visit7" = allData[, grep(".m60", colnames(allData), value = TRUE)],
                 # "visit8" = allData[, grep(".m72", colnames(allData), value = TRUE)]
                 #)
visitData <- cbind.data.frame(demogsAndOthersDf,visitData$visitbl,visitData$visit1,visitData$visit2, visitData$visit3, visitData$visit4)
#visitData = Reduce(function(x, y) merge(x, y, all=TRUE), list(demogsAndOthersDf,visitData$visitbl,visitData$visit1,visitData$visit2, visitData$visit3, visitData$visit4, visitData$visit5, visitData$visit6))

```

##impute the data#
```{r}
#colnames(visitData$visitbl) = sub("bl", "m00", colnames(visitData$visitbl)) 
# Create auxillary columns - group-wise and visit-wise
#visitData_aux = sapply(visitData, get_aux_all_groups)

#Impute value (visit wise)
#set.seed(123)
#visitData_imputed = imputedData = sapply(visitData_aux, function(x)missForest::missForest(x, ntree = 500)[1])


# set.seed(123)
# ##Make it as one data frame##
# imputedData <- cbind.data.frame(demogsAndOthersDf, visitData_imputed$visitbl.ximp, visitData_imputed$visit1.ximp,
#                             visitData_imputed$visit2.ximp, visitData_imputed$visit3.ximp, dxDf)
# imputedDataCp = imputedData
# colClean <- function(x){ colnames(x) <- gsub("m00", "bl", colnames(x)); x } 
# imputedData <- colClean(imputedData) 
# colClean <- function(x){ colnames(x) <- gsub("csf_aux", "csf.aux", colnames(x)); x } 
# imputedData <- colClean(imputedData) 
# colClean <- function(x){ colnames(x) <- gsub("brain_aux", "brain.aux", colnames(x)); x } 
# imputedData <- colClean(imputedData) 
# colClean <- function(x){ colnames(x) <- gsub("Cog_aux", "Cog.aux", colnames(x)); x } 
# imputedData <- colClean(imputedData) 
# colClean <- function(x){ colnames(x) <- gsub("m00", "bl", colnames(x)); x } 
# imputedData <- colClean(imputedData) 
# colClean <- function(x){ colnames(x) <- gsub("m00", "bl", colnames(x)); x } 
# imputedData <- colClean(imputedData) 
# colClean <- function(x){ colnames(x) <- gsub("csf_", "", colnames(x)); x } 
# imputedData <- colClean(imputedData) 
# colClean <- function(x){ colnames(x) <- gsub("brain_", "", colnames(x)); x } 
# imputedData <- colClean(imputedData) 
# colClean <- function(x){ colnames(x) <- gsub("Cog_", "", colnames(x)); x } 
# imputedData <- colClean(imputedData) 
# colClean <- function(x){ colnames(x) <- gsub("snp_", "", colnames(x)); x } 
# imputedData <- colClean(imputedData) 
# colClean <- function(x){ colnames(x) <- gsub("cortical_", "", colnames(x)); x } 
# imputedData <- colClean(imputedData) 
# colClean <- function(x){ colnames(x) <- gsub("_", ".", colnames(x)); x } 
# imputedData <- colClean(imputedData) 
# #setnames(imputedData, old = grep("rs[0-9]+", colnames(imputedData), value= TRUE), new = gsub("\\.bl","", grep("rs[0-9]+", colnames(imputedData), value= TRUE)))
# #setnames(imputedData, old = grep("Left|Right", colnames(imputedData), value= TRUE), new = gsub("\\.bl","", grep("Left|Right", colnames(imputedData), value= TRUE)))
#visitData_OBB = sapply(visitData_aux, function(x) missForest::missForest(x, ntree = 100)[2])
save.image("~/Documents/Documents_IT/ADNIVAMBN_paper_all/VAMBNForADNI/ADNIVAMBN.RData") 
```
