---
title: "adasSubscores"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = getwd())
```

## subset the data for adas sub-scores, adas 1 to adas 11

```{r cars}
#https://pdfs.semanticscholar.org/ed89/7aad02f24231123f0682fdf885aca5ff6d69.pdf
#adas_cogData = read.csv(paste(getwd(), '/AddNeuroMed/Data from Groupshare/ADAS-COG/data.csv', sep = ""))
load(paste(getwd(), "/ADNIMERGE - 23.10.2019/ADNIMERGE/data/adas.rdata", sep = ""))
adas_cogData = adas
##subset character columns except entity_id##
adasSubscores = grep("ADAS\\_[0-9]*$", colnames(adas_cogData), value = TRUE)
adasSubSet = adas_cogData[,c("RID",adasSubscores, "ADAS_COG_Total")]
##replacing Not_applicable with NA
adasSubSet[adasSubSet == "Not_applicable"] = NA
adasSubSet[adasSubSet == ""] = NA
```


## replacing the text in adas individual data with values 
```{r, echo = FALSE}
library(stringr)
##extract the columns with character##
charCols = colnames(adasSubSet[,sapply(adasSubSet, class) %in% c('factor', 'character')])
print(charCols)
charCols = setdiff(charCols, "entity_id")
adasSubSet[,charCols] <- lapply(adasSubSet[,charCols], as.character)
adas_cogVariables <- read_csv(paste(getwd(), "/AddNeuroMed/Data from Groupshare/ADAS-COG/variables.csv", sep =""))
adas_cogVariables$`categories:en` <- sapply(adas_cogVariables$`categories:en`, gsub, pattern = "=", replacement= ", ")
adas_cogVariables$`categories:en` <- sapply(adas_cogVariables$`categories:en`, gsub, pattern = ";", replacement= "\n")

##extract numeric values for adas scores from text
for(names in charCols){
  for(i in 1:nrow(adasSubSet)){
      print(is.na(adasSubSet[i,names]))
      if(is.na(adasSubSet[i, names]) == FALSE){
        if(adasSubSet[i, names] != ""){
       text <- adasSubSet[i, names]
       print("Text")
       print(text)
       meaningOfText = subset(adas_cogVariables$`categories:en`, adas_cogVariables$name == names)
       print(meaningOfText)
       loc <- str_locate(meaningOfText, paste(text, "\\,(.*|.*\n.*)\\([0-9]+\\)", sep = ""))
       print(loc)
       numValue <- substr(meaningOfText, loc[1,"start"], loc[1, "end"])
       print(numValue)
       numValue <- gsub(".*\\(", "", numValue)
       print(numValue)
       numValue <- (gsub("\\)", "", numValue))
       print(numValue)
       numValue <- as.numeric(as.character(numValue))
       adasSubSet[i,names]=numValue
      }
    }
  }
}
##converting all character columns to numeric
adasSubSet[,c(adasSubscores, "ADAS_COG_Total")] <- lapply(adasSubSet[,c(adasSubscores, "ADAS_COG_Total")], as.numeric)
```

##Validating of the generation of numeric scores for individual ADAS
```{r}
##checking if all features are missing for a patient
indexAllNAadas <- rowSums(is.na(adasSubSet)) == ncol(adasSubSet)-1
adasSubSet <- adasSubSet[which(indexAllNAadas == FALSE),]
indexAllNAadasFP <- rowSums(is.na(adasSubSet)) == (ncol(adasSubSet)-2)
adasSubSet <- adasSubSet[which(indexAllNAadasFP == FALSE),]
adasSubSet$ADAS_total_validate <- rowSums(adasSubSet[,adasSubscores], na.rm = TRUE)
idx <- sub('.*\\_', '', adasSubSet$ADAS_COG_Total) == sub('.*\\_', '', adasSubSet$ADAS_total_validate)
adasSubSet[which(idx == FALSE), "entity_id"]
```


##dividing the scores into 4 or 3 cateogaries, memory, language, praxis and concentration(not in this case,only in ADAS 13)
```{r}
##for AddNeuroMed
##memory: add ADAS 1, 4, 6, 7, 8 AND 9
adasSubSet$ADAS_Memory = rowSums(adasSubSet[,c("ADAS_1", "ADAS_4", "ADAS_6", "ADAS_7", "ADAS_8", "ADAS_9")])

##language: add ADAS 2, 10, 11
adasSubSet$ADAS_Language = rowSums(adasSubSet[,c("ADAS_2", "ADAS_10", "ADAS_11")])

##praxis: add ADAS 5, 3
adasSubSet$ADAS_Praxis = rowSums(adasSubSet[,c("ADAS_3", "ADAS_5")])
```

##save workspace
```{r, echo=FALSE}
save.image(paste(getwd(), "/Scripts/adasPreprocessing.RData", sep = ""))
save(adasSubSet, file = paste(getwd(), "/Scripts/adasSubSet.RData", sep = ""))
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
